Smoothing Algorithms for Variable Rate Models

### 2012-04-19 finaldraft_VRsmoothing.tex
Final draft.

\graphicspath{{./images/}{./images/finance/}{./images/tracking/}}

@@

\begin{figure}
\fbox{\parbox{\columnwidth}{
\begin{algorithmic}[1]
\linespread{1.5} \selectfont
  \STATE Run Rao-Blackwellised variable rate particle filter to approximate $p(\theta_{n}|y_{1:n})$ with weighted particles $\{\theta_{n}^{(i)}, w_{n}^{(i)}\}$ and $\{p(x_n|\theta_{n}^{(i)},y_{1:n})\}$ as normal distributions with moments $\{m_{n}^{(i)}\}$ and $\{P_{n}^{(i)}\}$. Store all results.
  \FOR{$i=1 \dots N_S$}
  	\STATE Initialise particle using $\theta^{(i)} \sim \sum_j w_N^{(j)} \delta_{\theta^{(j)}}(\theta)$.
    \STATE Initialise sufficient statistics $\tilde{m}_N^{(i)}$ and $\tilde{P}_N^{(i)}$ (see text).
    \FOR{$n=N-1 \dots 1$}
      \STATE Discard $\theta_{n}^{(i)}$.
      \STATE Predict state mean and covariance $\tilde{m}_n^{-(i)}$ and $\tilde{P}_n^{-(i)}$ using \ref{eq:backward_kf_predict_start} to \ref{eq:backward_kf_predict_stop}.
      \FOR{$j=1 \dots N_P$}
	      \STATE Calculate weight $\tilde{w}_n^{(j)}$ using (\ref{eq:rb-vrps_back_cond_weight2}).
      \ENDFOR
      \STATE Sample $\theta_{n}^{(i)} \sim \sum_j \tilde{w}_n^{(j)} \delta_{\theta_{n}^{(j)}}(\theta_{n})$.
      \STATE Join $\theta^{(i)} \gets \theta_{n}^{(i)} \cup \theta_{n}^{+(i)}$.
      \STATE Update state mean and covariance $\tilde{m}_n^{(i)}$ and $\tilde{P}_n^{(i)}$ using \ref{eq:backward_kf_update_start} to \ref{eq:backward_kf_update_stop}.
    \ENDFOR
  \ENDFOR
\end{algorithmic}
}}
\caption{Rao-Blackwellised Variable Rate Particle Smoother}
\label{alg:RBVRPS}
\end{figure}

@@

This model may be discretised at the observation times using standard methods (see e.g. \cite{Grewal2002}), leading to,
%
%Assuming observations of value only and Gaussian observation noise with standard deviation $\sigma_y^2$, the resulting discrete time dynamics are described by the following equations:% (see appendix \ref{app:model_discretisation}):
%
\begin{IEEEeqnarray}{rCl}
 \mathbf{x}_n &=& A \mathbf{x}_{n-1} + \mathbf{w}_n \\
 y_n          &=& C \mathbf{x}_{n} + v_n
\end{IEEEeqnarray}

@@

\begin{figure}[!t]
\centering
\includegraphics[width=0.9\columnwidth]{finance_unique_particles.pdf}
\caption{Mean number of unique particles in the filter-smoother (dashed) and smoother (solid) estimates.}
\label{fig:finance_unique_particles}
\end{figure}

@@

Solving these yields the following state equations (where $\Delta\ t = t - \tau_{K(t)}$):
%
\begin{IEEEeqnarray}{rCl}
\dot{s}(t) & = & \dot{s}_{K(t)} + a_{T,k} \Delta t \label{eq:2D_ICmodel_start} \\
\psi(t) & = & \psi_{K(t)} + \frac{a_{N,k}}{a_{T,k}} \log \left( \frac{\dot{s}(t)}{\dot{s}_{K(t)}} \right) \\
x(t) & = & x_{K(t)} + d_{X,K(t)} \Delta t \\
     \IEEEeqnarraymulticol{3}{l}{ \quad + \: \frac{ \dot{s}(t)^2 }{ 4 a_{T,k}^2 + a_{N,k}^2 } \left[  a_{N,k} \sin(\psi(t)) + 2 a_{T,k} \cos(\psi(t))  \right]} \nonumber \\
     \IEEEeqnarraymulticol{3}{l}{ \quad - \: \frac{\dot{s}_{K(t)}^2}{4 a_{T,k}^2 + a_{N,k}^2} \left[  a_{N,k} \sin(\psi_{K(t)}) + 2 a_{T,k} \cos(\psi_{K(t)})  \right]} \nonumber \\
y(t) & = & y_{K(t)} + d_{Y,K(t)} \Delta t \\
     \IEEEeqnarraymulticol{3}{l}{ \quad + \: \frac{ \dot{s}(t)^2 }{ 4 a_{T,k}^2 + a_{N,k}^2 } \left[ -a_{N,k} \cos(\psi(t)) + 2 a_{T,k} \sin(\psi(t))  \right]} \nonumber \\
     \IEEEeqnarraymulticol{3}{l}{ \quad - \: \frac{\dot{s}_{K(t)}^2}{4 a_{T,k}^2 + a_{N,k}^2} \left[  -a_{N,k} \cos(\psi_{K(t)}) + 2 a_{T,k} \sin(\psi_{K(t)})  \right]} \nonumber      .
\end{IEEEeqnarray}

@@

\begin{figure}[!t]
\centering
\includegraphics[width=0.9\columnwidth]{tracking_unique_particles.pdf}
\caption{Mean number of unique particles in the final filter (dashed) and smoother (solid) estimates.}
\label{fig:tracking_unique_particles}
\end{figure}

@@

%\newpage
\bibliographystyle{IEEEtran}
\bibliography{D:/pb404/Dropbox/PhD/OTbib}
\input{D:/pb404/Dropbox/PhD/Biographies/IEEE-Bunch.tex}
\input{D:/pb404/Dropbox/PhD/Biographies/IEEE-Godsill.tex}
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\graphicspath{{./images/}{./images/finance/}{./images/tracking/}}
%\interdisplaylinepenalty=1000

@@

\begin{figure}
\fbox{\parbox{\columnwidth}{
\begin{algorithmic}[1]
\linespread{1.5} \selectfont
  \STATE Run Rao-Blackwellised variable rate particle filter to approximate $p(\theta_{n}|y_{1:n})$ with weighted particles $\{\theta_{n}^{(i)}, w_{n}^{(i)}\}$ and $\{p(x_n|\theta_{n}^{(i)},y_{1:n})\}$ as normal distributions with moments $\{m_{n}^{(i)}\}$ and $\{P_{n}^{(i)}\}$. Store all results.
  \FOR{$i=1 \dots N_S$}
  	\STATE Initialise particle using $\theta^{(i)} \sim \sum_j w_N^{(j)} \delta_{\theta^{(j)}}(\theta)$.
    \STATE Initialise sufficient statistics $\tilde{m}_N^{(i)}$ and $\tilde{P}_N^{(i)}$ (see text).
    \FOR{$n=N-1 \dots 1$}
      \STATE Discard $\theta_{n}^{(i)}$.
      \STATE Predict state mean and covariance $\tilde{m}_n^{-(i)}$ and $\tilde{P}_n^{-(i)}$ using (\ref{eq:backward_kf_predict_start}) to (\ref{eq:backward_kf_predict_stop}).
      \FOR{$j=1 \dots N_P$}
	      \STATE Calculate weight $\tilde{w}_n^{(j)}$ using (\ref{eq:rb-vrps_back_cond_weight2}).
      \ENDFOR
      \STATE Sample $\theta_{n}^{(i)} \sim \sum_j \tilde{w}_n^{(j)} \delta_{\theta_{n}^{(j)}}(\theta_{n})$.
      \STATE Join $\theta^{(i)} \gets \theta_{n}^{(i)} \cup \theta_{n}^{+(i)}$.
      \STATE Update state mean and covariance $\tilde{m}_n^{(i)}$ and $\tilde{P}_n^{(i)}$ using \ref{eq:backward_kf_update_start} to \ref{eq:backward_kf_update_stop}.
    \ENDFOR
  \ENDFOR
\end{algorithmic}
}}
\caption{Rao-Blackwellised Variable Rate Particle Smoother}
\label{alg:RBVRPS}
\end{figure}

@@

This model may be discretised at the observation times using standard methods (see \cite{Godsill2007a,Christensen2012,Grewal2002}), leading to,
%
%Assuming observations of value only and Gaussian observation noise with standard deviation $\sigma_y^2$, the resulting discrete time dynamics are described by the following equations:% (see appendix \ref{app:model_discretisation}):
%
\begin{IEEEeqnarray}{rCl}
 \mathbf{x}_n &=& A \mathbf{x}_{n-1} + \mathbf{w}_n \\
 y_n          &=& C \mathbf{x}_{n} + v_n
\end{IEEEeqnarray}

@@

\begin{figure}[!t]
\centering
\includegraphics[width=0.75\columnwidth]{finance_unique_particles.pdf}
\caption{Mean number of unique particles in the filter-smoother (dashed) and smoother (solid) estimates.}
\label{fig:finance_unique_particles}
\end{figure}

@@

Solving these yields the following state equations (where $\Delta\ t = t - \tau_{K(t)}$):
%The target dynamics are governed by the following state equations \cite{Bunch2012a} (where $\Delta\ t = t - \tau_{K(t)}$):
%
\begin{IEEEeqnarray}{rCl}
\dot{s}(t) & = & \dot{s}_{K(t)} + a_{T,k} \Delta t \label{eq:2D_ICmodel_start} \\
\psi(t) & = & \psi_{K(t)} + \frac{a_{N,k}}{a_{T,k}} \log \left( \frac{\dot{s}(t)}{\dot{s}_{K(t)}} \right)
\end{IEEEeqnarray}
\begin{IEEEeqnarray}{rCl}
x(t) & = & x_{K(t)} + d_{X,K(t)} \Delta t \\
     \IEEEeqnarraymulticol{3}{l}{ \quad + \: \frac{ \dot{s}(t)^2 }{ 4 a_{T,k}^2 + a_{N,k}^2 } \left[  a_{N,k} \sin(\psi(t)) + 2 a_{T,k} \cos(\psi(t))  \right]} \nonumber \\
     \IEEEeqnarraymulticol{3}{l}{ \quad - \: \frac{\dot{s}_{K(t)}^2}{4 a_{T,k}^2 + a_{N,k}^2} \left[  a_{N,k} \sin(\psi_{K(t)}) + 2 a_{T,k} \cos(\psi_{K(t)})  \right]} \nonumber
\end{IEEEeqnarray}
\begin{IEEEeqnarray}{rCl}
y(t) & = & y_{K(t)} + d_{Y,K(t)} \Delta t \\
     \IEEEeqnarraymulticol{3}{l}{ \quad + \: \frac{ \dot{s}(t)^2 }{ 4 a_{T,k}^2 + a_{N,k}^2 } \left[ -a_{N,k} \cos(\psi(t)) + 2 a_{T,k} \sin(\psi(t))  \right]} \nonumber \\
     \IEEEeqnarraymulticol{3}{l}{ \quad - \: \frac{\dot{s}_{K(t)}^2}{4 a_{T,k}^2 + a_{N,k}^2} \left[  -a_{N,k} \cos(\psi_{K(t)}) + 2 a_{T,k} \sin(\psi_{K(t)})  \right]} \nonumber      .
\end{IEEEeqnarray}

@@

\begin{figure}[!t]
\centering
\includegraphics[width=0.75\columnwidth]{tracking_unique_particles.pdf}
\caption{Mean number of unique particles in the final filter (dashed) and smoother (solid) estimates.}
\label{fig:tracking_unique_particles}
\end{figure}

@@

%\newpage
\bibliographystyle{IEEEtran}
\bibliography{D:/pb404/Dropbox/PhD/Cleanbib}
\input{D:/pb404/Dropbox/PhD/Biographies/IEEE-Bunch.tex}
\input{D:/pb404/Dropbox/PhD/Biographies/IEEE-Godsill.tex}
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
